---
title: "Quant STM"
output: html_notebook
---

```{r}
library(here)
library(DBI)
library(RSQLite)
library(raster)
library(sf)
library(httr) # generic webservice package
library(tmap)
```

```{r}
#test
# mean function
mean_na <- function(x){
  x_mean <- mean(x, na.rm=TRUE)
  return(x_mean)
}

# connect
db <- dbConnect(RSQLite::SQLite(), 'C:/R_Drive/Data_Files/LPKS_Data/Data/Soil_Pedon_Databases/NRCS/KSSL/LabDataMart_4-17-23/ncss_labdata.sqlite')

dbListTables(db)

# list fields
# dbListFields(db, 'lab_pedon')
# dbListFields(db, 'lab_site')
# dbListFields(db, 'lab_physical_properties')
# dbListFields(db, 'lab_chemical_properties')
# dbListFields(db, 'lab_combine_nasis_ncss')
lab_layer <- dbGetQuery(db, "SELECT * from lab_layer;")
lab_site <- dbGetQuery(db, "SELECT * from lab_site;")
lab_pedon <- dbGetQuery(db, "SELECT * from lab_pedon;")
lab_phy <- dbGetQuery(db, "SELECT * from lab_physical_properties;")
lab_chem <- dbGetQuery(db, "SELECT * from lab_chemical_properties;")
lab_nasis <- dbGetQuery(db, "SELECT * from lab_combine_nasis_ncss;")

# series included in the Loamy Tableland ESD
Loamy_tableland <- c('Ulysses', 'Richfield', 'Keith', 'Kuma', 'Rosebud', 'Blackwood', 'Dawes', 'Mace', 'McConaughy', 'Norka', 'Satanta', 'Weld')

Loamy_tableland_nasis <- lab_nasis |> dplyr::filter(SSL_name %in% Loamy_tableland) |> dplyr::select(pedon_key, pedlabsampnum, peiid, samp_name,corr_name, SSL_name, lat=latitude_decimal_degrees, lon=longitude_decimal_degrees) 


phy_layer <- lab_phy |> dplyr::select(layer_key, texture_lab, clay_total, silt_total, sand_total, bulk_density_third_bar, aggregate_stability_05_2_mm, water_retention_third_bar, water_retention_15_bar, water_retention_field_state) |> dplyr::left_join(lab_layer |> dplyr::select(layer_key, labsampnum, hzn_top, hzn_bot, hzn_desgn, hzn_master, pedon_key, site_key), by="layer_key")

phy_chem_layer <- phy_layer |> dplyr::left_join(lab_chem |> dplyr::select(layer_key, ph_h2o,total_carbon_ncs, total_nitrogen_ncs, organic_carbon_walkley_black,estimated_organic_carbon, carbon_to_nitrogen_ratio), by="layer_key")


LT_phy_chem <- Loamy_tableland_nasis |> dplyr::left_join(phy_chem_layer , by=c("pedon_key")) |> dplyr::filter(!is.na(lat))

```

```{r}
#NASIS tables
# pedon  to siteobs: siteobsiidref==siteobsiid

pedon <- data.table::fread("C:/R_Drive/Data_Files/LPKS_Data/Data/Soil_Pedon_Databases/NRCS/NASIS/NASIS_APRIL_2017/CSV_files/pedon.csv",
    sep = "|", header = T)

siteobs <- data.table::fread("C:/R_Drive/Data_Files/LPKS_Data/Data/Soil_Pedon_Databases/NRCS/NASIS/NASIS_APRIL_2017/CSV_files/siteobs.csv",
    sep = "|", header = T)


nasis_pedon <- pedon |> dplyr::select(peiid, siteobsiidref, earthcovkind1, earthcovkind2, pedlabsampnum) |> dplyr::left_join(siteobs |> dplyr::select(siteobsiid, obsdate), by=c('siteobsiidref'='siteobsiid'))


#join to LT_phy_chem

LT_phy_chem_nasis <- LT_phy_chem |> dplyr::left_join(nasis_pedon, by="peiid") |> dplyr::distinct()
```


```{r}
LT_data_points <- st_as_sf(LT_phy_chem, coords = c("lon","lat"), crs = 4326)

# State/MLRA Boundary

#wfs_mlra <- 'https://services.arcgis.com/SXbDpmb7xQkk44JV/ArcGIS/rest/services/US_MLRA/FeatureServer'
url <- parse_url("https://services.arcgis.com/SXbDpmb7xQkk44JV/arcgis/rest/services")
url$path <- paste(url$path, "US_MLRA/FeatureServer/0/query", sep = "/")
url$query <- list(where = "1=1",
                  outFields = "*",
                  returnGeometry = "true",
                  f = "geojson")
request <- build_url(url)
mlra <- st_read(request)

#wfs_states <- 'https://services.arcgis.com/SXbDpmb7xQkk44JV/ArcGIS/rest/services/USA_States_Generalized/FeatureServer'
url <- parse_url("https://services.arcgis.com/SXbDpmb7xQkk44JV/arcgis/rest/services")
url$path <- paste(url$path, "USA_States_Generalized/FeatureServer/0/query", sep = "/")
url$query <- list(where = "1=1",
                  outFields = "*",
                  returnGeometry = "true",
                  f = "geojson")
request <- build_url(url)
states <- st_read(request)
states <- st_crop(states, mlra)


tmap_mode("view")
  tm_shape(states) +
    tm_borders() +
    tm_lines(lwd = "strokelwd", legend.lwd.show = FALSE) +
    tm_shape(LT_data_points) +
    tm_dots(col="#1B9E77", size=0.3) +
    
    #tm_scale_bar(position = c(0.06, 0.05)) +
    tm_add_legend('fill', 
	col = c( "#1B9E77"),
	border.col = "grey40",
	size = 1,
	labels = c('KSSL'),
	title="Potential Sampling Areas") +
    tm_layout(main.title = "Loamy Tableland ESD", bg.color = "white", legend.outside = TRUE) +
    tm_view(set.view = c(-99.22, 39.13,  4))

```

```{r}
library(raster)


nlcd_change <- raster::raster('C:/R_Drive/Data_Files/NRCS/NLCD_ESD/NLCD_2001_2019_change_index_L48_20210604_9ZdibphhsrPTxdgZKOjp.tiff')
Lt_nlcd_change <- raster::extract(nlcd_change, LT_data_points)

nlcd01 <- raster::raster('C:/R_Drive/Data_Files/NRCS/NLCD_ESD/NLCD_2001_Land_Cover_L48_20210604_9ZdibphhsrPTxdgZKOjp.tiff')
nlcd04 <- raster::raster('C:/R_Drive/Data_Files/NRCS/NLCD_ESD/NLCD_2004_Land_Cover_L48_20210604_9ZdibphhsrPTxdgZKOjp.tiff')
nlcd06 <- raster::raster('C:/R_Drive/Data_Files/NRCS/NLCD_ESD/NLCD_2006_Land_Cover_L48_20210604_9ZdibphhsrPTxdgZKOjp.tiff')
nlcd08 <- raster::raster('C:/R_Drive/Data_Files/NRCS/NLCD_ESD/NLCD_2008_Land_Cover_L48_20210604_9ZdibphhsrPTxdgZKOjp.tiff')
nlcd11 <- raster::raster('C:/R_Drive/Data_Files/NRCS/NLCD_ESD/NLCD_2011_Land_Cover_L48_20210604_9ZdibphhsrPTxdgZKOjp.tiff')
nlcd13 <- raster::raster('C:/R_Drive/Data_Files/NRCS/NLCD_ESD/NLCD_2013_Land_Cover_L48_20210604_9ZdibphhsrPTxdgZKOjp.tiff')
nlcd16 <- raster::raster('C:/R_Drive/Data_Files/NRCS/NLCD_ESD/NLCD_2016_Land_Cover_L48_20210604_9ZdibphhsrPTxdgZKOjp.tiff')
nlcd19 <- raster::raster('C:/R_Drive/Data_Files/NRCS/NLCD_ESD/NLCD_2019_Land_Cover_L48_20210604_9ZdibphhsrPTxdgZKOjp.tiff')
Lt_nlcd_01 <- raster::extract(nlcd01, LT_data_points)
Lt_nlcd_04 <- raster::extract(nlcd04, LT_data_points)
Lt_nlcd_06 <- raster::extract(nlcd06, LT_data_points)
Lt_nlcd_08 <- raster::extract(nlcd08, LT_data_points)
Lt_nlcd_11 <- raster::extract(nlcd11, LT_data_points)
Lt_nlcd_13 <- raster::extract(nlcd13, LT_data_points)
Lt_nlcd_16 <- raster::extract(nlcd16, LT_data_points)
Lt_nlcd_19 <- raster::extract(nlcd19, LT_data_points)

Lt_nlcd.df <- data.frame(Lt_nlcd_change, Lt_nlcd_01, Lt_nlcd_04, Lt_nlcd_06, Lt_nlcd_08, Lt_nlcd_11, Lt_nlcd_13, Lt_nlcd_16, Lt_nlcd_19) |> purrr::set_names('nlcd_change', 'nlcd01', 'nlcd04', 'nlcd06', 'nlcd08', 'nlcd11', 'nlcd13', 'nlcd16', 'nlcd19', )
Lt_nlcd.df <- Lt_nlcd.df |> dplyr::mutate(lc = dplyr::case_when(nlcd_change==1 & nlcd19==82 ~ 'Crops',
                                                         nlcd_change==1 & nlcd19==81 ~ 'Pasture',
                                                         nlcd_change==1 & nlcd19==71 ~ 'Grassland',
                                                         nlcd_change==7 & nlcd06==52 ~ 'Shrub-Crop07',
                                                         nlcd_change==7 & nlcd08==71 ~ 'Grassland-Crop09',
                                                         nlcd_change==7 & nlcd06==71 & nlcd08==82 ~ 'Grassland-Crop07',
                                                         TRUE ~ 'Fail'))

LT_data_points_nlcd_data <- dplyr::bind_cols(LT_phy_chem_nasis, Lt_nlcd.df)

LT_data_points_nlcd <- LT_data_points_nlcd_data |> dplyr::filter(!is.na(texture_lab), !hzn_bot < hzn_top, !hzn_bot == hzn_top)

aqp::depths(LT_data_points_nlcd) <-  pedon_key ~ hzn_top  + hzn_bot 
LT_data_points_nlcd.slab.r <- aqp::slab(LT_data_points_nlcd, fm = pedon_key ~ clay_total + silt_total + sand_total + bulk_density_third_bar + ph_h2o + organic_carbon_walkley_black + estimated_organic_carbon + aggregate_stability_05_2_mm, slab.structure=c(0,15), slab.fun=mean_na)
LT_data_points_nlcd.slab.r.1 <- LT_data_points_nlcd.slab.r |> dplyr::select(-c(contributing_fraction)) |> tidyr::pivot_wider(names_from=variable, values_from=value)
LT_data_points_nlcd.slab.r.1[ is.na(LT_data_points_nlcd.slab.r.1) ] <- NA
LT_data_points_nlcd.slab.r.1 <- LT_data_points_nlcd.slab.r.1 |> dplyr::rowwise() |> dplyr::mutate(soc = if_else(!is.na(organic_carbon_walkley_black), organic_carbon_walkley_black, estimated_organic_carbon))|> dplyr::mutate(soc_cl = soc_class(soc), ph_cl = ph_class(ph_h2o), db_cl = db_class(bulk_density_third_bar), txt_class = gettt(sand_total, silt_total, clay_total))


LT_data_points_complete <- LT_data_points_nlcd.slab.r.1 |> dplyr::filter(!is.na(soc_cl) & !is.na(db_cl) & !is.na(ph_cl) & !is.na(txt_class)) |> dplyr::select(pedon_key, top, bottom, soc_cl, ph_cl, db_cl, txt_class) |> dplyr::ungroup() |> dplyr::mutate(pedon_key = as.integer(pedon_key))

LT_data_points_nlcd_data_sub <- LT_data_points_nlcd_data |> dplyr::select(pedon_key, SSL_name, lat, lon, earthcovkind1,obsdate,lc) |> dplyr::distinct()

LT_data_points_complete <- LT_data_points_complete |> dplyr::left_join(LT_data_points_nlcd_data_sub, by="pedon_key")
```

#DSP4SH database
```{r}
library(DBI)
library(RSQLite)
library(dplyr, warn.conflicts = FALSE)
library(readxl)
library(readr)
library(here)

dsp4sh4 <- dbConnect(SQLite(), here("data/raw_data/dsp4sh4.db"))
dbListTables(dsp4sh4) #To check tables

pedon <- dbGetQuery(dsp4sh4, "SELECT * from pedon;")
coop_lab <- dbGetQuery(dsp4sh4, "SELECT * from cooplabmst;")
kssl_lab <- dbGetQuery(dsp4sh4, "SELECT * from kssllabmst;")
layerdesc <- dbGetQuery(dsp4sh4, "SELECT * from layerdescription;")
layerdesg <- dbGetQuery(dsp4sh4, "SELECT * from layerdesignation;")
plot <- dbGetQuery(dsp4sh4, "SELECT * from plotoverview;")
dspplotmgt <- dbGetQuery(dsp4sh4, "SELECT * from dspplotmgt;")

lab <- coop_lab |> dplyr::select(-c(DSP_Pedon)) |> dplyr::left_join(kssl_lab, by=c('KSSL_labsampnum'='natural_key'))
pedon_lab <- pedon |> dplyr::left_join(lab, by='DSP_Pedon_ID')
pedon_lab <- pedon_lab |> dplyr::left_join(dspplotmgt |> dplyr::select(-c(Soil)), by='DSP_Plot_ID')
pedon_lab <- pedon_lab |> dplyr::left_join(layerdesc |> dplyr::rename(field_clay_pct = Clay_pct), by=c('DSP_Pedon_ID'='DSP_Pedon_ID', 'DSP_sample_ID'='DSP_sample_ID'))
pedon_lab <- pedon_lab |> dplyr::left_join(layerdesg, by=c('DSP_Pedon_ID'='DSP_Pedon_ID', 'DSP_sample_ID'='DSP_sample_ID'))
pedon_lab <- pedon_lab |> dplyr::filter(Soil=="Keith")

pedon_lab <- pedon_lab |> dplyr::select(c(pedon_ID,DSP_Pedon_ID,DSP_Plot_ID,DSP_sample_ID,LU,till,trt, pedon_x,pedon_y,Soil,Bulk_Density,Water_Content,SOC_pct,KSSL_WSA,Yoder_AggStab_MWD,Soil_Respiration,Bglucosidase,Bglucosaminidase,AlkalinePhosphatase,AcidPhosphatase,Phosphodiesterase,POX_C,ACE,KSSL_labsampnum,horizon_designation,lay_depth_to_top, lay_depth_to_bottom,clay_tot_psa,silt_tot_psa,sand_tot_psa,co3_cly,tex_psda,adod,caco3,ph_cacl2,ph_h2o,CKMnO4,PNitroBGlu,ag_stab,c_tot_ncs,n_tot_ncs,s_tot_ncs,estimated_organic_C,Field_Texture,Coarse_Frag_volume,field_clay_pct,Color_Moist_Hue,Color_Moist_Value,Color_Moist_Chroma))


pedon_lab_points <- st_as_sf(pedon_lab, coords = c("pedon_x","pedon_y"), crs = 4326)

dsp_nlcd_change <- raster::extract(nlcd_change, pedon_lab_points)
dsp_nlcd_01 <- raster::extract(nlcd01, pedon_lab_points)
dsp_nlcd_04 <- raster::extract(nlcd04, pedon_lab_points)
dsp_nlcd_06 <- raster::extract(nlcd06, pedon_lab_points)
dsp_nlcd_08 <- raster::extract(nlcd08, pedon_lab_points)
dsp_nlcd_11 <- raster::extract(nlcd11, pedon_lab_points)
dsp_nlcd_13 <- raster::extract(nlcd13, pedon_lab_points)
dsp_nlcd_16 <- raster::extract(nlcd16, pedon_lab_points)
dsp_nlcd_19 <- raster::extract(nlcd19, pedon_lab_points)

dsp_nlcd.df <- data.frame(dsp_nlcd_change, dsp_nlcd_01, dsp_nlcd_04, dsp_nlcd_06, dsp_nlcd_08, dsp_nlcd_11, dsp_nlcd_13, dsp_nlcd_16, dsp_nlcd_19) |> purrr::set_names('nlcd_change', 'nlcd01', 'nlcd04', 'nlcd06', 'nlcd08', 'nlcd11', 'nlcd13', 'nlcd16', 'nlcd19', )
dsp_nlcd.df <- dsp_nlcd.df |> dplyr::mutate(lc = dplyr::case_when(nlcd_change==1 & nlcd19==82 ~ 'Crops',
                                                         nlcd_change==1 & nlcd19==81 ~ 'Pasture',
                                                         nlcd_change==1 & nlcd19==71 ~ 'Grassland',
                                                         nlcd_change==7 & nlcd06==52 ~ 'Shrub-Crop07',
                                                         nlcd_change==7 & nlcd08==71 ~ 'Grassland-Crop09',
                                                         nlcd_change==7 & nlcd06==71 & nlcd08==82 ~ 'Grassland-Crop07',
                                                         TRUE ~ 'Fail'))

dsp_data_points_nlcd_data <- dplyr::bind_cols(pedon_lab, dsp_nlcd.df)

aqp::depths(dsp_data_points_nlcd_data) <-  pedon_ID ~ lay_depth_to_top  + lay_depth_to_bottom 
dsp_data_points_nlcd.slab.r <- aqp::slab(dsp_data_points_nlcd_data, fm = pedon_ID ~ clay_tot_psa + silt_tot_psa + sand_tot_psa + Bulk_Density + ph_h2o + SOC_pct + KSSL_WSA, slab.structure=c(0,15), slab.fun=mean_na)
dsp_data_points_nlcd.slab.r.1 <- dsp_data_points_nlcd.slab.r |> dplyr::select(-c(contributing_fraction)) |> tidyr::pivot_wider(names_from=variable, values_from=value)
dsp_data_points_nlcd.slab.r.1[ is.na(dsp_data_points_nlcd.slab.r.1) ] <- NA
dsp_data_points_nlcd.slab.r.1 <- dsp_data_points_nlcd.slab.r.1 |> dplyr::rowwise() |> dplyr::mutate(soc_cl = soc_class(SOC_pct), ph_cl = ph_class(ph_h2o), db_cl = db_class(Bulk_Density), txt_class = gettt(sand_tot_psa, silt_tot_psa, clay_tot_psa))
dsp_data_points_nlcd.slab.r.1 <- dsp_data_points_nlcd.slab.r.1 |> dplyr::ungroup() |> dplyr::rename(pedon_key=pedon_ID)

dsp_data_points_complete <- dsp_data_points_nlcd.slab.r.1 |> dplyr::filter(!is.na(soc_cl) & !is.na(db_cl) & !is.na(ph_cl) & !is.na(txt_class)) |> dplyr::select(pedon_key, top, bottom, soc_cl, ph_cl, db_cl, txt_class) |> dplyr::ungroup() |> dplyr::mutate(pedon_key = as.integer(pedon_key))

dsp_data_points_nlcd <- dplyr::bind_cols(pedon_lab, dsp_nlcd.df)
dsp_data_points_nlcd_data_sub <- dsp_data_points_nlcd |> dplyr::select(pedon_key=pedon_ID, SSL_name=Soil, lat=pedon_y, lon=pedon_x, earthcovkind1=trt,lc) |> dplyr::distinct()
LT_data_points_nlcd_data_sub <- LT_data_points_nlcd_data |> dplyr::select(pedon_key, SSL_name, lat, lon, earthcovkind1,obsdate,lc) |> dplyr::distinct()
dsp_data_points_complete <- dsp_data_points_complete |> dplyr::left_join(dsp_data_points_nlcd_data_sub, by="pedon_key")

```

# combine dataframes
```{r}
stm_data <- bind_rows(dsp_data_points_complete, LT_data_points_complete |> dplyr::select(-c(obsdate)))

stm_data <- stm_data |> dplyr::mutate(soc_cl = factor(soc_cl, order=TRUE, levels=c('low', 'moderate', 'high', 'very high')), db_cl = factor(db_cl, order=TRUE, levels=c('very loose', 'loose', 'normal', 'hard', 'compact')), ph_cl = factor(ph_cl, order=TRUE, levels=c("Strongly acid", "Moderately acid", "Slightly acid", "Neutral", "Slightly alkaline", "Moderately alkaline")), txt_class = factor(txt_class, order=TRUE, levels=textures)) 

stm_data <- stm_data |> dplyr::mutate(lc= if_else(lc=="Fail", "Grassland", lc))


```

```{r}

library(bnclassify)
stm_data_sub <- stm_data |> dplyr::filter(lc=="Crops" | lc=="Grassland") |> dplyr::select(lc, txt_class, db_cl, ph_cl, soc_cl) |> dplyr::mutate(lc=factor(lc)) |> as.data.frame()
stm_data_sub <- stm_data |> dplyr::select(lc, txt_class, db_cl, ph_cl, soc_cl) |> dplyr::mutate(lc=factor(lc))

stm_data_sub <- stm_data_sub |> dplyr::mutate(txt_class = factor(txt_class, order=F), ph_cl = factor(ph_cl, order=F), soc_cl = factor(soc_cl, order=F), db_cl = factor(db_cl, order=F)) |> as.data.frame()

nb <- nb('lc', stm_data_sub) # Learn a naive Bayes structure
nb <- lp(nb, stm_data_sub, smooth = 1) # Learn parameters

cv(nb, stm_data_sub, k = 10) # 10-fold Cross-validation estimate of accuracy
#> [1] 0.8576045
head(predict(nb, stm_data_sub)) # Classify the entire data set
#> [1] unacc unacc unacc unacc unacc unacc
#> Levels: unacc acc good vgood
#> 
#> # Naive Bayes
nb <- nb('lc', stm_data_sub)
# ODE Chow-Liu with AIC score (penalized log-likelihood)
ode_cl_aic <- tan_cl('lc', stm_data_sub, score = 'aic')
# Semi-naive Bayes with forward sequential selection and joining (FSSJ) and
# 5-fold cross-validation
fssj <- fssj('lc', stm_data_sub, k = 5, epsilon = 0)

plot(ode_cl_aic)


nb <- lp(nb, stm_data_sub, smooth = 0.01)
awnb <- lp(nb, stm_data_sub, smooth = 0.01, awnb_trees = 10, awnb_bootstrap = 0.5)
manb <- lp(nb, stm_data_sub, smooth = 0.01, manb_prior = 0.5)
ode_cl_aic <- bnc('tan_cl', 'lc', stm_data_sub, smooth = 1, dag_args = list(score = 'aic'))
logLik(ode_cl_aic, stm_data_sub)
AIC(ode_cl_aic, stm_data_sub)
p <- predict(nb, stm_data_sub)
accuracy(p, stm_data_sub$lc)

set.seed(0)
cv(ode_cl_aic, stm_data_sub, k = 10)
cv(ode_cl_aic, stm_data_sub, k = 20, dag = FALSE, mean = FALSE)
cmi('txt_class', 'soc_cl', stm_data_sub, 'lc')



library(mlr)
ode_cl_aic_mlr <- as_mlr(ode_cl_aic, dag = TRUE, id = "ode_cl_aic")

# 5-fold cross-validation
rdesc = makeResampleDesc("CV", iters = 2)
# sequential floating forward search
ctrl = makeFeatSelControlSequential(method = "sfs", alpha = 0)
# Wrap ode_cl_aic_mlr with feature selection
ode_cl_aic_mlr_fs = makeFeatSelWrapper(ode_cl_aic_mlr, resampling = rdesc,
control = ctrl, show.info = FALSE)
t <- makeClassifTask(id = "stm_data_sub", data = stm_data_sub,
target = 'lc', fixup.data = "no", check.data = FALSE)

suppressWarnings(RNGversion("3.5.0"))
set.seed(0)
# Select features
mod <- train(ode_cl_aic_mlr_fs, task = t)
sfeats <- getFeatSelResult(mod)
sfeats

```


# Pixel /Index Value Description
1 No Change
2 Change from or to Water
3 Change from or to any of the four Urban classes (open space; low, medium, and high intensity)
4 Change from Herbaceous Wetland to Woody Wetland, or vice versa
5 Change from or to Herbaceous Wetland
6 Change from Cultivated Crops to Hay / Pasture, or vice versa
7 Change from or to Cultivated Crops
8 Change from or to Hay / Pasture
9 Persistent Grassland and Shrubland change. This change index attempts to identify changes to persistent Grassland and Shrubland areas, and to separate them from transitional shrubland areas such as regenerating forests.
10 Change from or to Barren
11 Change from or to any of the three Forest classes (Evergreen, Deciduous, and Mixed)
12 Change from or to Woody Wetland


# National Land Cover Database Class Legend and Description
Class\ Value Classification Description

## Water
11 Open Water- areas of open water, generally with less than 25% cover of vegetation
or soil.

12 Perennial Ice/Snow- areas characterized by a perennial cover of ice and/or snow,
generally greater than 25% of total cover.

## Developed
21 Developed, Open Space- areas with a mixture of some constructed materials, but
mostly vegetation in the form of lawn grasses. Impervious surfaces account for less
than 20% of total cover. These areas most commonly include large-lot single-family
housing units, parks, golf courses, and vegetation planted in developed settings for
recreation, erosion control, or aesthetic purposes.

22 Developed, Low Intensity- areas with a mixture of constructed materials and
vegetation. Impervious surfaces account for 20% to 49% percent of total cover.
These areas most commonly include single-family housing units.

23 Developed, Medium Intensity -areas with a mixture of constructed materials and
vegetation. Impervious surfaces account for 50% to 79% of the total cover. These
areas most commonly include single-family housing units.

24 Developed High Intensity-highly developed areas where people reside or work in
high numbers. Examples include apartment complexes, row houses and
commercial/industrial. Impervious surfaces account for 80% to 100% of the total
cover.

## Barren
31 Barren Land (Rock/Sand/Clay) - areas of bedrock, desert pavement, scarps, talus,
slides, volcanic material, glacial debris, sand dunes, strip mines, gravel pits and other
accumulations of earthen material. Generally, vegetation accounts for less than 15%
of total cover.

## Forest
41 Deciduous Forest- areas dominated by trees generally greater than 5 meters tall,
and greater than 20% of total vegetation cover. More than 75% of the tree species
shed foliage simultaneously in response to seasonal change.

42 Evergreen Forest- areas dominated by trees generally greater than 5 meters tall,
and greater than 20% of total vegetation cover. More than 75% of the tree species
maintain their leaves all year. Canopy is never without green foliage.

43 Mixed Forest- areas dominated by trees generally greater than 5 meters tall, and
greater than 20% of total vegetation cover. Neither deciduous nor evergreen species
are greater than 75% of total tree cover.

## Shrubland
51 Dwarf Scrub- Alaska only areas dominated by shrubs less than 20 centimeters tall
with shrub canopy typically greater than 20% of total vegetation. This type is often
co-associated with grasses, sedges, herbs, and non-vascular vegetation.

52 Shrub/Scrub- areas dominated by shrubs; less than 5 meters tall with shrub canopy
typically greater than 20% of total vegetation. This class includes true shrubs, young
trees in an early successional stage or trees stunted from environmental conditions.

## Herbaceous
71 Grassland/Herbaceous- areas dominated by gramanoid or herbaceous vegetation,
generally greater than 80% of total vegetation. These areas are not subject to
intensive management such as tilling, but can be utilized for grazing.

72 Sedge/Herbaceous- Alaska only areas dominated by sedges and forbs, generally
greater than 80% of total vegetation. This type can occur with significant other
grasses or other grass like plants, and includes sedge tundra, and sedge tussock
tundra.

73 Lichens- Alaska only areas dominated by fruticose or foliose lichens generally
greater than 80% of total vegetation.

74 Moss- Alaska only areas dominated by mosses, generally greater than 80% of total
vegetation.

## Planted/Cultivated
81 Pasture/Hay-areas of grasses, legumes, or grass-legume mixtures planted for
livestock grazing or the production of seed or hay crops, typically on a perennial
cycle. Pasture/hay vegetation accounts for greater than 20% of total vegetation.

82 Cultivated Crops -areas used for the production of annual crops, such as corn,
soybeans, vegetables, tobacco, and cotton, and also perennial woody crops such as
orchards and vineyards. Crop vegetation accounts for greater than 20% of total
vegetation. This class also includes all land being actively tilled.

## Wetlands
90 Woody Wetlands- areas where forest or shrubland vegetation accounts for greater
than 20% of vegetative cover and the soil or substrate is periodically saturated with
or covered with water.


# Denomination ph range
Ultra acid < 3.5
Extremely acid 3.5–4.4
Very strongly acid 4.5–5.0
Strongly acid 5.1–5.5
Moderately acid 5.6–6.0
Slightly acid 6.1–6.5
Neutral 6.6–7.3
Slightly alkaline 7.4–7.8
Moderately alkaline 7.9–8.4
Strongly alkaline 8.5–9.0
Very strongly alkaline > 9.0



# SOC classes (4 classes)
<!-- 0-1 -->
<!-- 1-2 -->
<!-- 2-3 -->
<!-- >3 -->



# Db classes
<!-- <1 = very loose -->
<!-- 1-1.2 = loose -->
<!-- 1.2-1.4 = normal -->
<!-- 1.4-1.6 = hard -->
<!-- >1.6 = compact -->



```{r}
textures <- c("Sand","Loamy sand","Sandy loam","Loam","Silt loam",
                "Silt","Sandy clay loam","Clay loam","Silty clay loam",
                "Sandy clay","Silty clay","Clay")

ph_class <- function(ph){
if(is.na(ph)){
  class = NA
}else if(ph <= 3.5){
  class = 'Ultra acid'
}else if(ph > 3.5 & ph <= 4.5){
  class = 'Extremely acid'
}else if(ph > 4.5 & ph <= 5){
  class = 'Very strongly acid'
}else if(ph > 5 & ph < 5.6){
  class = 'Strongly acid'
}else if(ph >= 5.6 & ph <= 6){
  class = 'Moderately acid'
}else if(ph > 6 & ph <= 6.5){
  class = 'Slightly acid'
}else if(ph > 6.5 & ph <= 7.3){
  class = 'Neutral'
}else if(ph > 7.3 & ph <= 7.8){
  class = 'Slightly alkaline'
}else if(ph > 7.8 & ph <= 8.4){
  class = 'Moderately alkaline'
}else if(ph > 8.4 & ph <= 9){
  class = 'Strongly alkaline'
}else if(ph > 9 & ph <= 9){
  class = 'Very strongly alkaline'
}else {class = NA}
return(class)
}

soc_class <- function(soc){
if(is.na(soc)){
  class = NA
}else if(soc <= 1){
  class = 'low'
}else if(soc > 1 & soc <= 2){
  class = 'moderate'
}else if(soc > 2 & soc <= 3){
  class = 'high'
}else if(soc > 3){
  class = 'very high'
}else {class = NA}
return(class)
}

db_class <- function(Db){
if(is.na(Db)){
  class = NA
}else if(Db <= 1){
  class = 'very loose'
}else if(Db > 1 & Db <= 1.2){
  class = 'loose'
}else if(Db > 1.2 & Db <= 1.4){
  class = 'normal'
}else if(Db > 1.4 & Db <= 1.6){
  class = 'hard'
}else if(Db > 1.6){
  class = 'compact'
}
return(class)
}

gettt <- function(sand, silt, clay){
  if(is.na(sand) | is.na(silt) | is.na(clay)){
    x = NA
  } else if((silt + 1.5 * clay) < 15){
    x = "Sand"
  } else if((silt + 1.5 * clay) >= 15 & (silt + 2.0 * clay) < 30){
    x = "Loamy sand"
  } else if((clay >= 7) & (clay <= 20) & (sand > 52) & ((silt + 2.0 * clay) >= 30)){
    x = "Sandy loam"
  } else if((clay < 7) & (silt < 50) & ((silt + 2.0 * clay) >= 30)){
    x = "Sandy loam"
  } else if((clay >= 7) & (clay <= 27) & (silt >= 28) & (silt < 50) & (sand <= 52)){
    x = "Loam"
  } else if(((silt >= 50) & (clay >= 12) & (clay < 27)) | ((silt >= 50) & (silt < 80) & (clay < 12))){
    x = "Silt loam"
  } else if((silt >= 80) & (clay < 12)){
    x = "Silt"
  } else if((clay >= 20) & (clay < 35) & (silt < 28) & (sand > 45)){
    x = "Sandy clay loam"
  } else if((clay >= 27) & (clay < 40) & (sand > 20) & (sand <= 45)){
    x = "Clay loam"
  } else if((clay >= 27) & (clay < 40) & (sand <= 20)){
    x = "Silty clay loam"
  } else if((clay >= 35) & (sand >= 45)){
    x = "Sandy clay"
  } else if((clay >= 40) & (silt >= 40)){
    x = "Silty clay"
  } else if((clay >= 40) & (sand <= 45) & (silt < 40)){
    x = "Clay"
  }
  return(x)
}

```

